---
title: "Reproducible Research"
author: "Luke Munyan"
date: "April 12, 2017"
output: html_document
---

Peer Graded Assignment: Course Project 1
=========================================

Read in the data and reformat the date column to be in a true date format

```{r}
activity<-read.csv("C://Users//Luke.Munyan//Documents//activity.csv", header=TRUE,stringsAsFactors = FALSE)
activity$date<-as.Date(activity$date, "%Y-%m-%d")
str(activity)
```

Next, we will calculate total steps each day.  We will make a histogram of the data to see what the most common step counts in a given day are.

```{r}
daily.steps<-aggregate(steps~date, data=activity, sum, na.rm=TRUE)
names(daily.steps)<-c("Date", "Steps")
```

```{r}
hist(daily.steps$Steps, main="Histogram of Steps Each Day", xlab="Steps Per Day", col="orange")
```

Now we will look at the mean and median number of steps per day.

```{r}
mean(daily.steps$Steps)
median(daily.steps$Steps)
```

What does the daily activity pattern look like?

```{r}
average_steps<-aggregate(steps~interval, data=activity, mean, na.rm=T)

names(average_steps)<-c("Interval", "Steps")
```

```{r ggplot, results="hide"}
require(ggplot2)
```

```{r}
ggplot(average_steps, aes(Interval, Steps))+geom_line(color="blue", size=.75)+ggtitle("Average Number of Steps at Each Given Interval Throughout the Day")
```

Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?

```{r}
max.int<-average_steps[average_steps$Steps==max(average_steps$Steps),]
max.int[,1]
```

How many missing values are there in the dataset?

```{r}
nrow(activity[is.na(activity$steps),])
```

We will fill the missing values with the overall average number of steps for the given interval the data is missing for. I will add a new row to the original data fram called "new steps" which will fill in the NA averages for that interval, and keep the actual (non-na) values that were already present in the steps column. 

```{r}
activity$newsteps<-ifelse(is.na(activity$steps)==TRUE, tapply(activity$steps,activity$interval, mean, na.rm=T), activity$steps)
```

Aggregate the data now that there are no missing values

```{r}
new.daily.steps<-aggregate(newsteps~date, data=activity, sum, na.rm=TRUE)
names(new.daily.steps)<-c("Date", "Steps")
```

Histogram with missing data filled in.

```{r}
hist(new.daily.steps$Steps, main="Histogram of Steps Each Day 2", xlab="Steps Per Day", col="purple")
```

Mean and Median

```{r}
mean(new.daily.steps$Steps)

median(new.daily.steps$Steps)
```

What is the difference between weekend and weekday activity patterns?
First we will create a new column/variable

```{r}
for(i in 1:nrow(activity)){
if(weekdays(activity$date[i])=="Saturday"){
    activity$week[i]<-"Weekend"
}else if(weekdays(activity$date[i])=="Sunday"){
    activity$week[i]<-"Weekend"
}else{activity$week[i]<-"Weekday"}
}
```

We will compare activity using a ggplot pannel plot after aggregating the data.

```{r}
new<-aggregate(newsteps~interval+week, data=activity, mean)
```

```{r}
ggplot(new, aes(interval, newsteps, color=week))+geom_line(size=.25, show.legend = F)+facet_grid(week~.)+xlab("Interval")+ylab("Average Steps")+ggtitle("Average Steps Throughout the Day by Weekday/Weekend")
```
